---
title: "Introduction to R"
author: "Lukáš 'hejtmy' Hejtmánek"
date: "21/07/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

Benefits
- It's free!!!
- It's multi platform
- It's actively developed
- It has wide community of package developpers, online support
- It has command line - interpretted language
- It can draw beautiful graphs
- It's easy unlike other programming languages
- It's made for numerical analysis (unlike python)
- It has elegant code (unlike spss) and is generally more powerfull
- It can deal with any data - parsing webpages, loading sql, loading data from google sheets etc.

Cons
- learning curve
- in the beginning, takes getting used to and takes longer than SPSS or excel
- limited visual 

## Demostrations
[Ebola data](https://gallery.shinyapps.io/Ebola-Dynamic/)
[Bus data](https://gallery.shinyapps.io/086-bus-dashboard/)
[Lego sets](https://shiny.rstudio.com/gallery/lego-set.html)
[Movie data](https://shiny.rstudio.com/gallery/movie-explorer.html)

## Your way around R studio
- Environemtn
- text editor
- console
- terminal
- History
- File manager
- Global options/project options


## Variables and functions
Examples of variables adn functions

```{r}
number <- -5
abs(number)
log(abs(number))
exp(number)

char <- "hello all"
toupper(char)
paste(char, "you in here")

vec <- 1:100
floor(vec)

```
Many of the basic functions can be found [here](statmethods.net/management/functions.html)

## Help in R and online
You can get help to any function in R by typing ?

```{r}
# ?read.csv
```

# Packages

Packages are packaged 

```{r}
library(ggplot2)
# install.packages("psych")
```

# Script example and breakdown

```{r}
library(ggplot2)
head(mpg) #function
str(mpg)
summary(mpg) #function
g <- ggplot(mpg, aes(class))
g + geom_bar()
```


```{r}
mpg <- ggplot2::mpg
# fitting simple anova model
cty_man_aov <- aov(cty ~ manufacturer, mpg)
summary(cty_man_aov)
```

```{r}
g <- ggplot(mpg, aes(manufacturer, cty))
g + geom_boxplot()
```

# Coding style and recommendation

Why coding style is important
- 

General recommendations

Coding style is described "here"
