---
title: "Vectors and vector subsetting"
author: "Lukáš 'hejtmy' Hejtmánek"
date: "21/07/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# vectors 

Basically anything in R is a vector. IN other languages, these may be called arrays (jacascript/c#) or lists(python). You can get length of a vector 

```{r}
num <- 1
length(num)
length("hello")
```

## Creating vectors
Most commonly you are going to be creating vectors by reading in data, but here are few important functions to generate your own dataset.

```{r}
vec <- c(1,2,3,4,5,6,7,8,9)
vec <- seq(1, 9, 1)
vec <- rep(1, 10)
vec <- 1:9
vec2 <- 10:19
vec_complete <- c(vec, vec2)
```

Tasks

- Create vector of 5 cities called cities with 5 EU capitals. It must include Prague
- Create vector temperatures of human temperature from 36.3 to 40.5 with 0.1 steps
- create vector roulette_numbers with numbers from 0 to 36
- Create vector called roulette_colors with "green" and then 18 times "red" and "black"
- create vector called weekdays with each day of the week 
- create vector called weekdays_10 which is days in 10 weeks consecutively 
- Crete vector of odd numbers call odd_numbers with numbers from 1 to 99
- Create vector of length 100 called true_false (T, F, T, F, T, F, T, F ...)

```{r}
cities <- c("Prague", "Madrid", "London", "Berlin", "Rome")
temperatures <- seq(36.3,40.5,0.1)
```

### Randomizing

```{r}
sample(1:10, 10)
sample(1:10, 10, replace=T)
```

Tasks
 - sample temperature of 5 people
 - sample 10 roulette numbers properly
 

```{r}
set.seed(1)
sample(1:10, 1)
```

seeding

Tasks
 - set seed to 666
 - sample temeprature of 10 people and save to subject_temperature
 - sample 10 roulette numbers and save to roulette_turns

## Accessing vectors

### Indexing by number
R has 1 based indexing, meaning the first position in a vector has

```{r}
vec <- 1:9
vec[1]
vec[5]
vec[vec[2]]
vec[-5]
vec[-c(1:4)]
```

Tasks
- select monday from the weekdays vector
- select all mondays from teh weekdays_10
- select prague from the cities vector
- select all false from the true_false vector
- select all "black numbers from the roulette numbers
- what colours did we draw in roulette turns?

### Indexing by Logical

```{r}
vec_bool <- c(T,T,F,F,T,T)
vec_num <- 1:6
vec_num[vec_bool]
```

R automatically repeats vectors if you run out of them
```{r}
bool <- c(TRUE, FALSE)
vec <- 1:100
vec[bool]
vec[FALSE]
```

Taks
- use the true_false vector to create the odd_numbers vector
- select every other day in the weekdays_10 vector


### Indexing by condition
`Which` function

```{r}
which(cities == "Prague")
i_prague <- which(cities == "Prague")
cities[i_prague]
cities[-i_prague]

cities[cities != "Prague"]
```

`%in` function

```{r}
fav_cities <- c("Prague", "London")
which(cities %in% fav_cities)
```

Task
- what are indices of "monday" in your weekdays_10 vector?
- what are indices of either weekend day in your weekdays_10 vector?
- what are indices of "black" in your roulettes_colors vector?

```{r}

```

## Selecting by condition

Most common and easiest to read
```{r}
temp_above_37 <- temperatures > 37
temperatures[temp_above_37]
temperatures[temperatures > 37]
```

Tasks
- how many possible temperatures are above 37.5C?
- 

## Vector functions

### Vector math

Adding vectors togehter
```{r}

```

Tasks

Most R functions work on vectors

```{r}
vec <- 1:100

mean(vec)
median(vec)

```

tasks
- what is the mean and median temeprature of people in the 



# Final exam

Each of these should be written in code on a single line

1. Create a vector of years between 1924 until 2019 by one year
2. Create a vector of chinese zodiac for each year, beginning with rat, repeating 10 times (120 years)
3. What years are for "horse"?
4. What sign was year 1987?
5. What is the sum of years for the "rooster"?
6. What is the median year for the "dragon"
7. What is the 3rd_year of a goat modulo 7?

```{r}
signs <- rep(c("rat", "ox", "tiger", "rabbit", "dragon", "snake", "horse", "goat", "monkey", "rooster", "dog", "pig"), 10)
years <- 1924:2019
i_years <- years - 1923

years[i_years[signs=="horse"]] #1
signs[years==1987] #2
sum(years[i_years[signs=="rooster"]], na.rm=T)
median(years[i_years[signs=="dragon"]], na.rm=T)
years[i_years[signs=="goat"]][3] %% 7
```

