---
title: "Vector statistics"
author: "Lukáš 'hejtmy' Hejtmánek"
date: "21/07/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Vector statistics

Let's create data first

Task:
- set seed of 666
- create a vector of length 100 likes_blue of randomized TRUE and FALSE
- create a vector of length 100 favorite_color of randomized ("red", "green", "yellow", "blue")
- create a vector of length 100 gender of randomized "male", "female"
- create a vector of length 100 height with a mean of 170 and SD 10 (use function rnorm)
- create a vector of length 100 salary with a mean of 29000 and SD 5000
- create a vector of length 100 hours_slept wchih is an uniform sample from 5-11

```{r}
set.seed(666)
likes_blue <- sample(c(T,F), 100, replace=T)
favorite_color <- sample(c("red", "green", "yellow", "blue"), 100, replace=T)
gender <- sample(c("female", "male"), 100, replace=T)
height <- rnorm(100, 170, 10)
salary <- rnorm(100, 29000 , 5000)
hours_slept <- sample(5:11, 100, replace=T)
```

Summaries of the variables
```{r}
table(likes_blue)
table(gender)
summary(height)
summary(hours_slept)
```

## Basic statistical tests
```{r}
cor(salary, height)
t.test(salary ~ gender)
lm_salary_height <- lm(salary ~ height)
summary(lm_salary_height)
summary(aov(salary ~ favorite_color))
```


## Subsetting things
Tasks
- what is the mean hours slept for people who like blue?
- what is the distributions of fravorite colors for people with salary above 35k?
- does the correlation change between salary adn height between women and men?


## Plotting things

### Histogram

```{r}
hist(height)
hist(height, breaks = 20)
hist(height, breaks = 50,  col="blue")
d <- density(mtcars$mpg) # returns the density data 
plot(d) # plots the results
```

### Bar plot


```{r}
counts <- table(favorite_color)
barplot(counts, main="Favorite color")
counts <- table(gender, favorite_color)
barplot(counts, main="Favorite color by gender", beside=TRUE, legend=rownames(counts))
```

### Scatter plot
```{r}
plot(height, salary)
abline(lm(salary ~ height))
title("Regression of salary on height")
```

Tasks
- plot salary as an effect of hours slept
- plot two separate plots (one after another) for salary as an effect of height separately for males and females 

### Boxplots
```{r}
boxplot(height)
```

## Creating new variables
When we use certain "search" often, we might as well save it in an variable

Task
- create new variable height_180 is "height > 180"
- create new variable dislike_red wchih is for people who didn't say they like "red"
- create new variable female_blue which is true for females who like blue color
```{r}
height_180 <- height > 180
dislike_red <- favorite_color != "red"
female_blue <- gender=="female" & favorite_color == "blue"
```
