---
title: "Loading data"
author: "Lukáš 'hejtmy' Hejtmánek"
date: "22/07/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Getting and setting folders

All loading and saving is going to be relative to your current "working directory". You can get working directory by `getwd()` and set it by `setwd('path to new')`

```{r}
#getwd()
#setwd
```

# Loading dataset without any issues

Just check that you have correct separator and it should be fine
```{r}
df <- read.table("../../../data/loading-examples/example-row.txt", sep=";", header=T)
head(df, 2)
str(df)
```

## Checking that everything is fine

```{r}
head(df, 3)
tail(df, 3)
summary(df)
str(df)
```

If you wanna view or edit the contents, you can use Rstudios `View(df)` function (capital V) or `edit(df)`. If you want to save edited df, don't forget to use `df <- edit(df)`

You can use `edit` to even create data.frames.
```{r}
mydata <- data.frame(age=numeric(0), gender=character(0), weight=numeric(0))
#mydata <- edit(mydata)
```

## Loading options

separator
header
decimal-point
encoding

### Strings vs factors

Factors are integers with character labels
```{r}
levels(df$favorite_color)
as.numeric(df$favorite_color[1:10])
as.character(df$favorite_color[1:10])
```

if you don't wan't that conversion by default, you can load the table with `stringsAsFactors = FALSE` or you can convert them later with `as.character`.

[More on the diferences here](https://swcarpentry.github.io/r-novice-inflammation/12-supp-factors/)

# Loading data from other sources

## Web

```{r}
df <- read.table("https://raw.githubusercontent.com/lab-code/R-for-non-programmers/cebex/data/loading-examples/example-row.txt", sep=";", header=T)
head(df, 2)
```

more on direct loading from the web [www.r-bloggers.com/getting-data-from-an-online-source](https://www.r-bloggers.com/getting-data-from-an-online-source/)

## Reading excel
```{r}
#install.packages("readxl)
df <- readxl::read_xlsx("../../../data/loading-examples/data.xlsx", sheet="Sheet2")
head(df, 2)
```

# Loading different datasets with different headers/commas/issues

```{r}
list.files("../../../data/loading-examples/")
```

### Final exam

- read in all files from the loading-examples dataset into table df
- check that you load it well by running the following code

```{r}
df_correct <- read.table("../../../data/loading-examples/example-row.txt", sep=";", header=T)
all(sapply(df, class)==c("integer","factor","factor","numeric","numeric","integer"))
all(df==df_correct)
```

[Complete tutorial on getting data into R](https://www.datacamp.com/community/tutorials/r-data-import-tutorial)
