---
title: "Data preparation"
author: "Lukáš 'hejtmy' Hejtmánek"
date: "22/07/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Donwload dataset about [student alcohol comsumtion](https://www.kaggle.com/uciml/student-alcohol-consumption/)


# Accessing data frames

Columns are basically just list fields, so can be accessed with `$` or with a column number

```{r}
df_math <- read.table("../../../data/student-alcohol-consumption/student-mat.csv", sep = ",", header=T, stringsAsFactors = F)
df_portugese <- read.table("../../../data/student-alcohol-consumption/student-por.csv", sep = ",", header=T, stringsAsFactors = F)
```

```{r}
df_math$school[1]
```

Data.frame uses [row, column] notation. So you can access first cell of thrid column as `df[1,3]`. 
If you want entire columns, you can do it in multiple ways

```{r, results='hide'}
df_math[, 3]
df_math[, "school"]
df_math$school
df_math[, c("school", "age")]
df_math[, c(1,5)]
df_math[, c(1:3, 8:10)]
```

# Merging datasets


Let's have a loot at the data first

Tasks
- do both have same column names?
- are there any obvious differences between the datasets?

## cbind
Cbind allows us to add new variable/columns to existing dataframe. It expects bound dataframes to have the same number of rows and be ordered appropriately. IN case, each si ordered differently, we need to bind them using IDs with `merge`

```{r}
df_test <- data.frame(id=1:100, gender=sample(c("M", "F"), 50, replace = T))
df_test_2 <- data.frame(test_result = rnorm(50, 100, 15))
df_all <- cbind(df_test, df_test_2)
colnames(df_all)
```

## rbind
rbind (row bind) allow us to append new rows to a data.frame

```{r}
df_test <- data.frame(id=1:100, gender=sample(c("M", "F"), 50, replace = T), test_result = rnorm(50, 100, 15))
nrow(df_test)
df_test_2 <- data.frame(id=101:200, gender=sample(c("M", "F"), 50, replace = T), test_result = rnorm(50, 100, 15))
df_all <- rbind(df_test, df_test_2)
nrow(df_all)
```
Applying on our student data

```{r}
df_students <- rbind(df_portugese, df_math)
```

## Removing duplicit data

By combinding these two datasets, we have created multiple logs for several students who are doing both portugese and math. We can find them using the duplicated function

```{r}
match_columns <- c("school","sex","age","address","famsize","Pstatus","Medu","Fedu","Mjob","Fjob",
                   "reason","nursery","internet","higher", "famsup", "freetime", "goout", "romantic",
                   "activities", "health", "famrel", "guardian", "traveltime","schoolsup")
sum(duplicated(df_students[, match_columns]))
```

## merge

Merge allows us to match data based on given id

```{r}
df_test <- data.frame(id=1:100, gender=sample(c("M", "F"), 50, replace = T))
df_test_2 <- data.frame(id=100:1, test_result = rnorm(50, 100, 15))
head(merge(df_test, df_test_2, by="id"))
```

By default merge only produces combination of both datasets, dropping any rows whcih are not matched in both datasets

```{r}
df_students <- merge(df_portugese, df_math, by=match_columns)
print(nrow(df_students))
```
if we want to keep even those which are not found in both we can do "left join" or "right join"

```{r}
df_students <- merge(df_portugese, df_math, by=match_columns, all = T)
print(nrow(df_students))
colnames(df_students)
```
And if we want to keep things tidy, we can rename the added suffixes
```{r}
df_students <- merge(df_portugese, df_math, by=match_columns, all = T, suffixes = c(".portugese", ".math"))
print(nrow(df_students))
colnames(df_students)
```

# Recoding variables
```{r}
str(df_students)
#address - student's home address type (binary: 'U' - urban or 'R' - rural)
df_students$address <- dplyr::recode(df_students$address, "R"="rural", "U"="urban")
table(df_students$address)
```

Tasks
- check that all binary columns (nursery, internet ...) only contain binary values and no missing data
- recode all binary columns to TRUE and FALSE

```{r}
df_students$famsup <- df_students$famsup == "yes"
df_students$schoolsup <- df_students$schoolsup == "yes"
```

## Using factors
Factors have an advantage over strings of allowing us to order them. Some of the fields int eh data are more likely to be factors than numeric values:

```{r}
df_students$traveltime  <- factor(df_students$traveltime , levels = 1:4, 
                                  labels = c("<15 min", "15-30 min","30-60 min",">60 min"), ordered = T)
min(df_students$traveltime)
```

Tasks
- factor studytime
- factor Medu, Fedu

```{r}

```

# New variables
Adding new variables is generally easy

```{r}
df_test <- data.frame(id=1:100, test_1=rnorm(100,100,15), test_2 = rnorm(100,100,15))
df_test$mean_test <- (df_test$test_1+df_test$test_2)/2
```

Tasks:
- create variable large_family to either true if famsize is GT3 and false if LE3
- Create any_support variable which is true if either famsup or schoolsup is true
- create variable both_courses which indicates if the student is taking both courses
- create variable either_higher_ed which is true if either father or mother education is higher
- create variable relationships which is a sum of famrel and a 3 if romantic is yes
- create variable total_failures which is a sum of math and portugeese failures (because of NA values, you will need `rowSums` function)
- create variable total_perfomance for math and portugeese which is a sum of G1-G3

```{r}
df_students$large_family <- df_students$famsize == "GT3"
df_students$any_support <- df_students$famsup | df_students$schoolsup
df_students$relationships <- df_students$famrel + 3*(df_students$romantic == "yes")
df_students$total_failures <- rowSums(df_students[,c("failures.portugese", "failures.math")], na.rm = T)
df_students$both_courses <- !(is.na(df_students$studytime.math) | is.na(df_students$studytime.portugese))
#df_students$parent_higher_ed <- df_students$
```

# Saving the data

saving of pure data is done with the write.table function

important arguments are:
- sep
- row.names
- col.names
- dec

RECOMMENDATION: DON'T save row names, use ";" as a separator

```{r}
#
```